---
title: "Fantasy Football - Post Draft"
author: "Robert Conacher"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE, echo = TRUE)
#Load Libraries
library(readxl)   
library(tidyverse)
library(rvest)
library(forcats)
library(latexpdf)
library(kableExtra)
library(scales)
library(treemap)
library(ggridges)
library(gt)
library(RColorBrewer)
```

```{r include = FALSE}
#Import Data
FF <- read_excel("FF-2020.xlsx", sheet = "FF")
mydata <- FF

#Load Team Data
Team <- read_excel("FF-2020.xlsx", sheet = "Teams")

#Clean Data
mydata$team <- factor(mydata$team)
mydata$position <- factor(mydata$position)
mydata$tier <- factor(mydata$tier)

#Clean Team Data
Team$Div <- factor(Team$Div)
Team$Conf <- factor(Team$Conf)
Team$TeamName <- factor(Team$TeamName)
Team$Byes <- factor(Team$Byes)

# Add Conf, Div & Full Team Names to dataset
mydata <- inner_join(mydata, Team, by = "team")


#Load Draft Data
Draft <- read_excel("Draft.xlsx", sheet = "Draft")

#Cleran Data
Draft$FF_Team <- as.factor(Draft$FF_Team)
Draft$D_round <- as.integer(Draft$D_round)
Draft$D_pick <- as.integer(Draft$D_pick)
Draft$adquire <- as.factor(Draft$adquire)
Draft$draft_pos <- as.integer(Draft$draft_pos)

# Join two data frames into one
FF <- full_join(Draft, mydata, by = c('players' = 'player')) %>%
  filter(FF_Team != " ")

#Impute missing points
# Asign to impute to show imputed or not
  FF <- FF %>%
  mutate(impute = if_else(is.na(points), 1,0))
  
# Store averages by position 
average_pts_pos <- FF %>%
  filter(!is.na(points)) %>%
  group_by(position) %>%
  summarize(mean(points))
average_pts_pos


FF <- FF %>%
  group_by(position) %>%
  mutate(mean_pts = if_else(is.na(points), mean(points,na.rm=TRUE), points))
  

FF %>%
  filter(impute == 1) %>%
  select(position, players, position, impute, points, mean_pts)

```


## Keepers

```{r Keeper1, include = FALSE}
FF %>%
  filter(FF_Team == "LA Rams") %>%
  ggplot(aes(x = position, y = mean_pts, color = Div, size = points)) + 
  geom_point() +
  facet_wrap(~Conf)
```

The 2020 season is about to start and the keepers are in and the draft is compl;eted. How did everyone do?
Well that's a slightly more complicated question to answer. 

There are many projections missing:

  1. impacts to teams and players due to the number of players not playing the season due to COVID-19
  2. lots of new rookies that have not been league tested
  3. Free agency on players that do not have a home this season 
  4. players returning from lengthy suspensions

Any project points that were missing due to the above were imputed with the average points for each position.

Total projections for all teams are slightly miss leading due to above issues. Also, some teams went very heavy on a specific position (such as Denver and Flutie Flakes - securing 4 Quarterbacks).

At the end, I have summarized for each Owner by their average projected points per position. This then allows Owners to do some comparisons to judge for themselves how well their draft / draft + keepers did against the other Owners.

Enjoy, and have a great season!

## Keepers Listing
```{r keeper1, echo = FALSE}
#KLeepers
#FF %>%
#  filter(adquire == "Keeper") %>%
#  select(FF_Team, players, position, mean_pts) %>%
#  group_by(FF_Team, mean_pts) %>%
#  mutate(FP = seq_len(n())) %>%
#  arrange(FF_Team, FP, desc(mean_pts), players) %>%
#  kbl() %>%
#  kable_paper("hover", full_width = F)



FF %>%
  filter(adquire == "Keeper") %>%
  select(FF_Team, players, position, mean_pts) %>%
  group_by(FF_Team) %>%
  mutate(FF_Team = fct_reorder(FF_Team, mean_pts)) %>%
  arrange(FF_Team, desc(mean_pts), players) %>%
  mutate(Keep = row_number()) %>%
  
  gt(rowname_col = "Keep", groupname_col = "FF_Team") %>%
  tab_header(
    title = md("**2020 Keepers**"),
    subtitle = md("Sorted by FF_Team & mean_pts")
  ) %>%
  tab_source_note(md(""))




```


## Keepers Chart
```{r keep2, echo = FALSE}
# Team Keepers
FF %>%
  filter(adquire == "Keeper") %>%
  select(FF_Team, players, position, mean_pts) %>%
  ggplot(aes(x = mean_pts , y = fct_reorder(players, mean_pts))) +
  geom_col() +
  scale_x_continuous(limits = c(0, 500)) +
  facet_wrap(~ FF_Team, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```


## Draft Listing
```{r draft1, echo = FALSE}
#Draft
#FF %>%
#  filter(adquire == "Draft") %>%
#  select(FF_Team, players, position, mean_pts) %>%
#  arrange(FF_Team, desc(mean_pts), players) %>%
#  kbl() %>%
#  kable_paper("hover", full_width = F)


FF %>%
  filter(adquire == "Draft") %>%
  select(FF_Team, players, position, mean_pts) %>%
  group_by(FF_Team) %>%
  mutate(FF_Team = fct_reorder(FF_Team, mean_pts)) %>%
  arrange(FF_Team, desc(mean_pts), players) %>%
  mutate(Keep = row_number()) %>%
  
  gt(rowname_col = "Draft", groupname_col = "FF_Team") %>%
  tab_header(
    title = md("**2020 Draft**"),
    subtitle = md("Sorted by FF_Team & mean_pts")
  ) %>%
  tab_source_note(md(""))

```


## Draft Chart
```{r draft2, echo = FALSE}
# Team Keepers
FF %>%
  filter(adquire == "Draft") %>%
  filter(!is.na(mean_pts)) %>%
  select(FF_Team, players, position, mean_pts) %>%
  ggplot(aes(x = mean_pts , fct_reorder(players, mean_pts))) +
  geom_col() +
  scale_x_continuous(limits = c(0, 500)) +
  facet_wrap(~ FF_Team, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



## Treeemap1
```{r treemap1, echo = FALSE}
# Treemap

treemap(FF,
            index=c("FF_Team","players"),
            vSize="mean_pts",
            type="index",
            title="2020 NFL FF Team Distributions (Based on Players and Projected Points)"
            ) 
```


## Treemap2

```{r treemap2, echo = FALSE}
#Treemap 2
FF %>%
   filter_all(any_vars(! is.na(.))) %>%
treemap(FF, #Your data frame object
        index=c("FF_Team", "position"),  #A list of your categorical variables
        vSize = "mean_pts",  #This is your quantitative variable
        type="index", #Type sets the organization and color scheme of your treemap
        palette = "Greens",  #Select your color palette from the RColorBrewer presets or make your own.
        title="2020 NFL FF Team Distributions (Based on Positions and Projected Points)", #Customize your title
        fontsize.title = 12 #Change the font size of the title
        )
```

```{r treetop3, echo = FALSE}
#Treemap 3
FF %>%
   filter_all(any_vars(! is.na(.))) %>%
treemap(FF, #Your data frame object
        index=c("position"),  #A list of your categorical variables
        vSize = "mean_pts",  #This is your quantitative variable
        type="index", #Type sets the organization and color scheme of your treemap
        palette = "Reds",  #Select your color palette from the RColorBrewer presets or make your own.
        title="2020 NFL Distributions (Based on Positions)", #Customize your title
        fontsize.title = 12 #Change the font size of the title
        )
```

```{r treetop4, echo = FALSE}
#Treemap 4
FF %>%
   filter_all(any_vars(! is.na(.))) %>%
treemap(FF, #Your data frame object
        index=c("Conf", "position"),  #A list of your categorical variables
        vSize = "mean_pts",  #This is your quantitative variable
        type="index", #Type sets the organization and color scheme of your treemap
        palette = "Reds",  #Select your color palette from the RColorBrewer presets or make your own.
        title="2020 NFL Distributions (Based on Positions within Conf)", #Customize your title
        fontsize.title = 12 #Change the font size of the title
        )
```

```{r treetop5, echo = FALSE}
#Treemap 5
FF %>%
   filter_all(any_vars(! is.na(.))) %>%
treemap(FF, #Your data frame object
        index=c("Conf", "Div", "position"),  #A list of your categorical variables
        vSize = "mean_pts",  #This is your quantitative variable
        type="index", #Type sets the organization and color scheme of your treemap
        palette = "Reds",  #Select your color palette from the RColorBrewer presets or make your own.
        title="2020 NFL Distributions (Based on Positions within Conf and Division)", #Customize your title
        fontsize.title = 12 #Change the font size of the title
        )
```


## Post Analysis

```{r last, echo = FALSE}
 Total_points <- FF %>%
  drop_na(points) %>%
  select(FF_Team, D_round, D_pick, players, position, mean_pts) %>%
  group_by(FF_Team) %>%
  summarise(Total_pts = sum(mean_pts)) %>%
  mutate(FF_Team = fct_reorder(FF_Team, desc(Total_pts))) %>%
  arrange(FF_Team, Total_pts)

# Total Projected Points post Draft by FF_Team   
 ggplot(Total_points, aes(FF_Team, Total_pts)) + 
    geom_point()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
    Total_points %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
 
  aa <- FF %>%
  select(FF_Team, D_round, D_pick, players, position, mean_pts) %>%
  drop_na(mean_pts) %>%
  group_by(FF_Team, position) %>%
  summarise(Mean_Pts_Pos = mean(mean_pts)) %>%
  mutate(FF_Team = fct_reorder(FF_Team, desc(Mean_Pts_Pos))) %>%
  arrange(FF_Team, position, Mean_Pts_Pos)
 
  aa %>%
    pivot_wider(names_from = "position", values_from = "Mean_Pts_Pos") 
  
  FF %>%
    ggplot(aes(x = position, y = mean_pts)) + 
    geom_boxplot() +
    facet_wrap(~ FF_Team)

FF %>%
  filter(!is.na(mean_pts)) %>%
  ggplot(aes(x = mean_pts, y = position)) +
  geom_density_ridges()+
  labs(title = "Point Distibution by Position for whole FF League")
  

#FF %>%
#  filter(!is.na(mean_pts)) %>%
#  filter(position == "DB") %>%
#  group_by(FF_Team) %>%
#  mutate(FF_Team = fct_reorder(FF_Team, mean_pts, mean)) %>%
#  ggplot(aes(x = FF_Team, y = mean_pts)) +
#  geom_point(aes(color = position)) +
#  geom_boxplot() +
#  geom_jitter(position = "jitter") +
#  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
#  facet_wrap(~ position, scales = "free") + 
#  coord_flip()
 
```

```{r Top5_pos, echo = FALSE}
# Who had the top 5 player at each position?
Top_5 <- FF %>%
    filter(!is.na(mean_pts)) %>%
    group_by(position) %>%
    mutate(position = fct_reorder(position, mean_pts)) %>% 
    top_n(n = 5) %>%
    select(position, mean_pts, players, adquire, FF_Team) %>%
    arrange(position, desc(mean_pts), players, adquire, FF_Team) 

Top_5 <- data.frame(Top_5)    
  
Top_5 %>%
  gt(rowname_col = "players", groupname_col = "position") %>%
  tab_header(
    title = md("**2020 FF - Top5 for Each Positions"),
    subtitle = md("Sorted by Position & Players")
  ) %>%
  tab_source_note(md(""))


```

```{r Top5_all, echo = TRUE}
# Who had the most top 5 regardless of position?    
Top_5 %>%
  group_by(FF_Team) %>%
  count(FF_Team, sort = TRUE)
```
```{r}
Top5_cnt <- Top_5 %>%
  count(FF_Team) 

ggplot(Top5_cnt, aes(x = fct_reorder(FF_Team, n), y = n)) +
  geom_col() +
  coord_flip() + 
  labs(title = "How many players were kept or drafted that are top 5 players",
        subtitle = "(Based on Projected Points)") +
  ylab("Total Number of top 5 players") +
  xlab("Fantasy Team - Owner")
```


## Summary

- Most of the top 5 for each position were keepers (77%) & draft (23%)
- 1 teams (Bishop Bucs) had 0 player in the top 5 positions
- 1 team (Mellonville) did not draft DL position
- Denver and Flutie Flakes - securing 4 Quarterbacks

